---
title: "Collaborative Research Project - Assignment 3"
bibliography:
date: "24 October 2014"
output: pdf_document
number_sections: yes
highlight: zenburn
authors: Meilin Moellenkamp and Nicolas Rosemberg
toc: yes
---


repmis::LoadandCite(pkgs, file = 'RpackageCitations.bi
Note: Use a file name that is different from your literature BibTeX file!
b')


```{r, echo=FALSE, include=FALSE}

### 1. Setting Working Directory & Loading the data

# error=FALSE
# Setting the Working Directory
setwd("/Users/Meilin/Desktop/Collaborative Social Data/CollaborativeResearchProject")
# setwd("/Users/Nico/Documents/Hertie/Social science data analysis/CollaborativeResearchProject")


# Loading the merged file from the Working directory
Merged <- read.csv(file="MergedData")
```

```{r, echo=FALSE, include=FALSE}

### 2. Load Required Packages

# install.packages("RJOSONIO")  
library(RJSONIO)
# install.packages("WDI")  
library(WDI)
# install.packages("dplyr")  
library(dplyr) 
# install.packages("tidyr")  
library(tidyr)
# install.packages("httr")  
library(httr) 
# install.packages("dplyr")  
library(dplyr)
# install.packages("XML")  
library(XML)
#install.packages("plyr")
library(plyr)
# install.packages("Amelia")  
library(Amelia) 
#install.packages("XLConnect")
library(XLConnect)    
# install.packages("countrycode")
library("countrycode")
# install.packages ("ggplot2")
library(ggplot2)
# # install.packages ("magrittr") ???
library(magrittr)
# install.packages ("fmsb")
library(fmsb)
# install.packages ("car")
library(car)

### 3. Check if Variables are coded as numeric
str(Merged)
# Code variables as numeric
Merged$Incidence <- as.numeric(Merged$Incidence)
Merged$DPT <- as.numeric(Merged$DPT)
Merged$Measles <- as.numeric(Merged$Measles)

```

# 1. Descriptive Statistics

For the discriptive statistics scatterplots and histograms will be shown in order to understand the distribution of the variables.

The histogram of the dependent variable (Figure 1) shows that the incidence rates are not normally distributed but strongly skewed to the left and only few incidence rates are higher than 1. 

```{r, echo=FALSE}

### 1. Plotting the dependent variable

# Histogram of dependent variable
Histdep <- hist(Merged$Incidence, breaks= seq(0,5,by=0.1), xlab="HIV Incidence Rate", main="Figure 1: Incidence Rate")
axis(side=1, at=seq(0,5, 0.5), labels=seq(0,5,0.5))
axis(side=2, at=seq(0,800,100), labels=seq(0,800,100))
```
