---
title: "Collaborative Research Project - Assignment 3"
bibliography:
date: "24 October 2014"
output: html_document
number_sections: yes
highlight: zenburn
authors: Meilin Moellenkamp and Nicolas Rosemberg
toc: yes
---

To automate the bibtex packages file:
repmis::LoadandCite(pkgs, file = 'RpackageCitations.bi
Note: Use a file name that is different from your literature BibTeX file!
b')

```{r, echo=FALSE, include=FALSE}

### 1. Setting Working Directory & Loading the data

# error=FALSE
# Setting the Working Directory
setwd("/Users/Meilin/Desktop/Collaborative Social Data/CollaborativeResearchProject")
# setwd("/Users/Nico/Documents/Hertie/Social science data analysis/CollaborativeResearchProject")


# Loading the merged file from the Working directory
Merged <- read.csv(file="MergedData")
```


```{r, echo=FALSE, include=FALSE}

### 2. Load Required Packages

# install.packages("RJOSONIO")  
library(RJSONIO)
# install.packages("WDI")  
library(WDI)
# install.packages("dplyr")  
library(dplyr) 
# install.packages("tidyr")  
library(tidyr)
# install.packages("httr")  
library(httr) 
# install.packages("dplyr")  
library(dplyr)
# install.packages("XML")  
library(XML)
#install.packages("plyr")
library(plyr)
# install.packages("Amelia")  
library(Amelia) 
#install.packages("XLConnect")
library(XLConnect)    
# install.packages("countrycode")
library("countrycode")
# install.packages ("ggplot2")
library(ggplot2)
# # install.packages ("magrittr") ???
library(magrittr)
# install.packages ("fmsb")
library(fmsb)
# install.packages ("car")
library(car)
# install.packages("stargazer")
library(stargazer)


### 3. Check if Variables are coded as numeric
str(Merged)
# Code variables as numeric
Merged$Incidence <- as.numeric(Merged$Incidence)
Merged$DPT <- as.numeric(Merged$DPT)
Merged$Measles <- as.numeric(Merged$Measles)

```

# 1. Descriptive Statistics

For the discriptive statistics scatterplots and histograms will be shown in order to understand the distribution of the variables.

The histogram of the dependent variable (Figure 1) shows that the incidence rates are not normally distributed but strongly skewed to the left and only few incidence rates are higher than 1. 

```{r, echo=FALSE}

### 1. Plotting the dependent variable

# Histogram of dependent variable
Histdep <- hist(Merged$Incidence, breaks= seq(0,5,by=0.1), xlab="HIV Incidence Rate", main="Figure 1: Incidence Rate")
axis(side=1, at=seq(0,5, 0.5), labels=seq(0,5,0.5))
axis(side=2, at=seq(0,800,100), labels=seq(0,800,100))

```

In most regions of the world HIV incidences decreased between the period of 2000 to 2015 (see Figure 2).

```{r, echo=FALSE}
# Look at incidence over time - seee the general trend
IncT <- ggplot(aes(x = year, y = Incidence), data = Merged) + geom_point(position = "jitter") + theme_bw() + geom_smooth()
options(digits=0)
IncT + ggtitle("Figure 2: Incidence Rate over Time") + theme(plot.title = element_text(lineheight=3, face="bold"))

ggsave("fig2.png", width=5, height=5, dpi=100)

```

When plotting the incidence rates per country (Figure 3) the countries with high incidence rates can be identified and the ranges of the observations per country are shown.

``` {r, echo=FALSE}
# Look at incidence rate per country
IncC <- ggplot2::ggplot(Merged, aes(Incidence, country)) + geom_point() + theme_bw()
IncC + ggtitle("Figure 3: Incidence Rate per Country") + theme(plot.title = element_text(lineheight=3, face="bold"))

ggsave("fig3.png", width=8, height=10, dpi=100)

```

As our research question is investigating why MDG 6.A is not being reached by some countries, we are not only interested in the general HIV incidence rate but also in the decrease in the incidence rate from 2000 to 2015. As stated in the introduction Target 6.A of the MDGs specifies that countries should "have halted by 2015 and begun to reverse the spread of HIV/AIDS"" (United Nations (2014)). 

For this purpose, the dependent variable was lagged by one period and the difference between the lag and the current year was calculated. Further, a dummy variable was created assigning a value of zero for those observations where the incidence rate decreased compared to the previous year or stayed the same (countries reaching MDG 6.A) and a value of one was assigned to those observations where the incidence rate increased (countries not reaching MDG 6.A). Figure 4 shows the direction of the change in the incidence rate compared to the previous year by country.

```{r, echo=FALSE}
# Look at difference in incidence rate per country between one year to the next
IncClag <- ggplot2::ggplot(Merged, aes(IncidenceDif, country)) + geom_point() + theme_bw()
IncClag + ggtitle("Figure 4: Change in Incidence Rate compared to Previous Years by Country") + theme(plot.title = element_text(lineheight=3, face="bold"))

ggsave("fig4.png", width=8, height=10, dpi=100)
```

Scatterplots were used for each category of the Dahlgrehn model in order to see whether the variables are skewed or multicollinear (Figures 5, 6 & 7).

```{r, echo=FALSE}
### 2. Plotting the independent variables

# Scatterplot of variables for socio-economic, cultural and environmental conditions
scatterplotMatrix(~ Incidence + + GDP + GDPpc + Rural + CO2 + LifeExpect + Population, 
                  transform=TRUE, data=Merged)

# Scatterplot of variables for living and working conditions 
scatterplotMatrix(~ Incidence + Water + Sanitation + Primary + Unemploym + HCexpendpc,
                  transform=TRUE, data=Merged)

# Scatterplot of variables for individual lifestyle factors
scatterplotMatrix(~ Incidence + DPT + Measles, transform=TRUE, data=Merged)

```


####################################################################################
########################### INFERENTIAL STATISTICS #################################
####################################################################################

As can be seen from the Scatterplots in the descriptive statistics most of the variables are not normally distributed. Further, the variables all have different scales. Therefore, the independent variables were logged for enabling comparisons in the inferential statistics part.

```{r, echo=FALSE, include=FALSE}
# Logging the independent variables for better comparability
Merged$lGDP <- log(Merged$GDP)
Merged$lGDPpc <- log(Merged$GDPpc)
Merged$lRural <- log(Merged$Rural)
Merged$lCO2 <- log(Merged$CO2)
Merged$lHCexpend <- log(Merged$HCexpend)
Merged$lWater <- log(Merged$Water)
Merged$lSanitation <- log(Merged$Sanitation)
Merged$lUnemploym <- log(Merged$Unemploym)
Merged$lPrimary <- log(Merged$Primary)
Merged$lHCexpendpc <- log(Merged$HCexpendpc)
Merged$lFemUnempl <- log(Merged$FemUnempl)
Merged$lFemSchool <- log(Merged$FemSchool)
Merged$lLifeExpect <- log(Merged$LifeExpect)
Merged$lDPT <- log(Merged$DPT)
Merged$lMeasles <- log(Merged$Measles)
```

As the dependent variable is coded as a dummy, being 0 if MDG 6.A was fulfilled and 1 for countries that don't fulfill MDG 6.A logistic regressions are used for the inferential statistics part. As Odds and Odds ratios are difficult to present to an audience that is not only consisting of statistical experts predicted probabilities are calculated following the logistic regressions. The interpretation of the results will only focus on the predicted probabilities.

```{r, echo=FALSE, include=FALSE}
# MODEL 1 - Running a general logistic regression using all independent variables
L1 <- glm(DDif ~ lGDP + lGDPpc + lRural + lCO2 + lHCexpend + lWater + lSanitation + lUnemploym + lPrimary + lHCexpendpc + lFemUnempl + lFemSchool + lLifeExpect + lDPT + lMeasles,
          data=Merged, family = 'binomial')
summary(L1)
```

```{r, echo=FALSE}
# Testing for multicollinearity
vif(L1)
```

The test for variance inflation factors showed that in our first logistic regression model six variables showed high multicollinearity and had a heigher variance inflation than the threshold of 10. We tested the multicollinearity between the variables and found that there was high multicollinearity between the GDP and GDP per capital, Unemployment and Female unemployment, Primary education and female schooling. Therefore, we excluded one of these variables for each group based on their explanatory strength for our research question, namely GDP.

```{r, echo=FALSE, include=FALSE}
# MODEL 2 - GDP
# Running the regression without multicollinear variables
L2 <- glm(DDif ~ lGDP + lRural + lCO2 + lHCexpend + lWater + lSanitation + lFemUnempl + lFemSchool + lLifeExpect + lDPT + lMeasles,
          data=Merged, family = 'binomial')

summary(L2)
```

```{r, echo=FALSE} 
# Testing for multicollinearity
vif(L2)
```

```{r, echo=FALSE} 
# MODEL 3 - GDPpc
L3 <- glm(DDif ~ lGDPpc + lRural + lCO2 + lHCexpend + lWater + lSanitation + lFemUnempl + lFemSchool + lLifeExpect + lDPT + lMeasles,
          data=Merged, family = 'binomial')
summary(L3)
# Testing for multicollinearity
vif(L3)

# Looking at confidence intervals
confint(L3)

# Testing for multicollinearity
bptest(L3)
coeftest(L3,vcov=hccm(L3))
```


```{r, echo=FALSE, include=FALSE}
# MODEL 4 - Including a new variable with FemSchool by quintiles

# Creating a new variable with FemSchool by quintiles
Merged$QFemSchool[Merged$lFemSchool<=4.53] <-1
Merged$QFemSchool[Merged$lFemSchool>4.53 & Merged$lFemSchool<=4.622] <-2
Merged$QFemSchool[Merged$lFemSchool>4.622 & Merged$lFemSchool<=4.697] <-3
Merged$QFemSchool[Merged$lFemSchool>4.697] <-4

# LOGISTIC REGRESSION with FemSchool by quintiles:
L4 <-  glm(DDif ~ lGDPpc + lRural + lCO2 + lHCexpend + lWater + lSanitation + 
            lFemUnempl + lLifeExpect + lDPT + lMeasles + as.factor(QFemSchool),
            data=Merged, family = 'binomial')
summary(L4)

# Look at Residual vs. Fitted Plot
plot(L4)

# PREDICTED PROBABILITIES:
#Regressing the model on FemSchool fixing the other indendent variables at Uganda's mean

# Looking at Uganda's average values
Uganda <- subset(Merged,(country=="Uganda"))
summary(Uganda)
summary(Merged)

# PREDICTED PROBABILITIES
fitted_L4 <- with(Merged,
                  data.frame(lGDPpc = 7.003,
                             lRural = 4.461,
                             lCO2 = -2.52,
                             lHCexpend = 2.125,
                             lLifeExpect = 3.98,
                             lWater = 4.18,
                             lSanitation = 3.46,
                             lDPT = 4.199,
                             lMeasles = 4.235,
                             lFemUnempl = 1.2576,
                             QFemSchool = factor(1:4)))
fitted_L4

# Calculating the predicted probabilities
fitted_L4_final <- predict(L4, newdata = fitted_L4, type = 'response')
fitted_L4_final

# MODEL 5 - with the new independent variable without the interaction
# LOGISTIC REGRESSION:
L5 <- glm(DDif ~ lGDPpc + lRural + lCO2 + lHCexpend + lWater + lSanitation + 
            lFemUnempl * lFemSchool + lLifeExpect + lDPT + 
            lMeasles,data=Merged, family = 'binomial')
 
summary(L5)

# Using the anova function to evaluate the effect on the deviance of adding the interaction term 
anova(L3,L5)
anova(L5,test="Chisq")

fitted_L5 <- with(Merged,
                  data.frame(lGDPpc = 7.003,
                             lRural = 4.461,
                             lCO2 = -2.52,
                             lHCexpend = 2.125,
                             lLifeExpect = 3.98,
                             lWater = 4.18,
                             lSanitation = 3.46,
                             lDPT = 4.199,
                             lMeasles = 4.235,
                             lFemUnempl = 1.2576,
                             QFemSchool = factor(1:4)))
fitted_L5

# Calculating the predicted probabilities
fitted_L5_final <- predict(L4, newdata = fitted_L5, type = 'response')
fitted_L5_final


```

Exporting the Predicted Probabilities with Stargazer
```{r, echo=FALSE, type = 'latex', header = FALSE }
# Saving the Regression Outputs with Stargazer

labels <- c('(Intercept)', 'lGDPpc', 'lRural', 'lCO2',
            'lHCexpend', 'lLifeExpect', 'lLifeExpect', 'lWater', 'lSanitation',
            'lDPT', 'lMeasles', 'lFemUnempl', '1st Quartile FemSchool', '2nd Quartile FemSchool',
            '3rd Quartile FemSchool', '4th Quartile FemSchool')


stargazer::stargazer(L4, covariate.labels = labels,
               title = 'Logistic Regression Estimates of HIV Incidence according to 
               Female Schooling', digits = 2, type = 'latex', header = FALSE)

```
If you want to automatically generate tables from regression model output objects, texreg is a good package to turn to. First estimate your models:

L1 <- glm(admit ~ gre,
              data = Admission, family = 'binomial')

L2 <- glm(admit ~ gre + gpa,
              data = Admission, family = 'binomial')

L3 <- glm(admit ~ gre + gpa + as.factor(rank),
              data = Admission, family = 'binomial')
Then use the stargazer function to create a results table. For PDFs set type = 'latex'. There are many stylistic modifications you can make with this function.
??????????????????????????????
# Create cleaner covariate labels
labels <- c('(Intercept)', 'GRE Score', 'GPA Score', '2nd Ranked School', '3rd Ranked School',
            '4th Ranked School')

stargazer::stargazer(L1, L2, L3, covariate.labels = la title = 'Logistic Regression Estimates of Grad Sch digits = 2, type = 'latex', header = FALSE)