---
title: "Collaborative Research Project - Assignment 3"
bibliography:
- Literature.bib
- Packages.bib
date: "24 October 2014"
output: html_document
number_sections: yes
highlight: zenburn
authors: Meilin Moellenkamp and Nicolas Rosemberg
toc: yes
---


repmis::LoadandCite(pkgs, file = 'RpackageCitations.bi
Note: Use a file name that is different from your literature BibTeX file!
b')


```{r, echo=FALSE, include=FALSE}

### 1. Setting Working Directory & Loading the data

# error=FALSE
# Setting the Working Directory
setwd("/Users/Meilin/Desktop/Collaborative Social Data/CollaborativeResearchProject")
# setwd("/Users/Nico/Documents/Hertie/Social science data analysis/CollaborativeResearchProject")


# Loading the merged file from the Working directory
Merged <- read.csv(file="MergedData")
```

```{r, echo=FALSE, include=FALSE}

### 2. Load Required Packages

# install.packages("RJOSONIO")  
library(RJSONIO)
# install.packages("WDI")  
library(WDI)
# install.packages("dplyr")  
library(dplyr) 
# install.packages("tidyr")  
library(tidyr)
# install.packages("httr")  
library(httr) 
# install.packages("dplyr")  
library(dplyr)
# install.packages("XML")  
library(XML)
#install.packages("plyr")
library(plyr)
# install.packages("Amelia")  
library(Amelia) 
#install.packages("XLConnect")
library(XLConnect)    
# install.packages("countrycode")
library("countrycode")
# install.packages ("ggplot2")
library(ggplot2)
# # install.packages ("magrittr") ???
library(magrittr)
# install.packages ("fmsb")
library(fmsb)
# install.packages ("car")
library(car)

### 3. Check if Variables are coded as numeric
str(Merged)
# Code variables as numeric
Merged$Incidence <- as.numeric(Merged$Incidence)
Merged$DPT <- as.numeric(Merged$DPT)
Merged$Measles <- as.numeric(Merged$Measles)

```

# 1. Descriptive Statistics

For the discriptive statistics scatterplots and histograms will be shown in order to understand the distribution of the variables.

The histogram of the dependent variable (Figure 1) shows that the incidence rates are not normally distributed but strongly skewed to the left and only few incidence rates are higher than 1. 

```{r, echo=FALSE}

### 1. Plotting the dependent variable

# Histogram of dependent variable
Histdep <- hist(Merged$Incidence, breaks= seq(0,5,by=0.1), xlab="HIV Incidence Rate", main="Figure 1: Incidence Rate")
axis(side=1, at=seq(0,5, 0.5), labels=seq(0,5,0.5))
axis(side=2, at=seq(0,800,100), labels=seq(0,800,100))
```

In most regions of the world HIV incidences decreased between the period of 2000 to 2015 (see Figure 2).

```{r, echo=FALSE}
# Look at incidence over time - seee the general trend
IncT <- ggplot(aes(x = year, y = Incidence), data = Merged) + geom_point(position = "jitter") + theme_bw() + geom_smooth()
options(digits=0)
IncT + ggtitle("Figure 2: Incidence Rate over Time") + theme(plot.title = element_text(lineheight=3, face="bold"))
```

When plotting the incidence rates per country (Figure 3) the countries with high incidence rates can be identified and the ranges of the observations per country are shown.

``` {r, echo=FALSE}
# Look at incidence rate per country
IncC <- ggplot2::ggplot(Merged, aes(Incidence, country)) + geom_point(position = "jitter") + theme_bw()
IncC + ggtitle("Figure 3: Incidence Rate per Country") + theme(plot.title = element_text(lineheight=3, face="bold"))

#opts_chunk$set(comment=NA, fig.width=10, fig.height=10)) ????

```

As our research question is investigating why MDG 6.A is not being reached by some countries, we are not only interested in the general HIV incidence rate but also in the decrease in the incidence rate from 2000 to 2015. As stated in the introduction Target 6.A of the MDGs specifies that countries should "have halted by 2015 and begun to reverse the spread of HIV/AIDS"" (United Nations (2014)). 

For this purpose, the dependent variable was lagged by one period and the difference between the lag and the current year was calculated. Further, a dummy variable was created assigning a value of zero for those observations where the incidence rate decreased compared to the previous year or stayed the same and a value of one was assigned to those observations where the incidence rate increased. Figure 4 shows the direction of the change in the incidence rate compared to the previous year by country.

```{r, echo=FALSE}
# Look at difference in incidence rate per country between one year to the next
IncClag <- ggplot2::ggplot(Merged, aes(IncidenceDif, country)) + geom_point(position = "jitter") + theme_bw()
IncClag + ggtitle("Figure 4: Change in Incidence Rate compared to Previous Years by Country") + theme(plot.title = element_text(lineheight=3, face="bold"))
```

Scatterplots were used for each category of the Dahlgrehn model in order to see whether the variables are skewed or multicollinear (Figures 5, 6 & 7).

```{r, echo=FALSE}
### 2. Plotting the independent variables

# Scatterplot of variables for socio-economic, cultural and environmental conditions
scatterplotMatrix(~ Incidence + + GDP + GDPpc + Rural + CO2 + LifeExpect + Population, 
                  transform=TRUE, data=Merged)
IncClag + ggtitle("Figure 5: Scatterplot of variables for socio-economic, cultural and environmental conditions") + theme(plot.title = element_text(lineheight=3, face="bold"))

# Scatterplot of variables for living and working conditions 
scatterplotMatrix(~ Incidence + Water + Sanitation + Primary + Unemploym + HCexpendpc,
                  transform=TRUE, data=Merged)
IncClag + ggtitle("Figure 6: Scatterplot of variables for living and working conditions") + theme(plot.title = element_text(lineheight=3, face="bold"))

# Scatterplot of variables for individual lifestyle factors
scatterplotMatrix(~ Incidence + DPT + Measles, transform=TRUE, data=Merged)
IncClag + ggtitle("Figure 7: Scatterplot of variables for individual lifestyle factors") + theme(plot.title = element_text(lineheight=3, face="bold"))
```
